### 2020-01-28 wuhan.tex
stuff

\begin{flushleft}{
	\Large
	\textbf\newline{
		Assessing uncertainties associated with early estimates of the basic reproduction number during the novel coronavirus (2019-nCoV) outbreak
	}
}
\newline
\\
Sang Woo Park\textsuperscript{1}
Jonathan Dushoff\textsuperscript{2,3}
\\

@@

As the novel coronavirus (2019-nCoV) continues to spread,
researchers are rushing to publish their estimates of the basic 
reproductive number $\mathcal R_0$.
The basic reproductive number (i.e., the 
average number of secondary cases generated 
by a primary case in a fully susceptible population) is of particular interest 
because it allows prediction about the final size of an epidemic.
While their efforts are valuable, their analyses rely on several
assumptions that could immediately affect their estimates of $\mathcal R_0$ and
the associated uncertainties in their estimates.
Here, we illustrate the degree to which their assumptions can affect
their estimates and stress few principles that needs to be taken into
consideration.

@@

Early in an outbreak, $\mathcal R_0$ cannot be estimated directly,
whereas the exponential growth rate $r$ can be estimated reliably.
Given estimates of the exponential growth rate $r$ and the distribution $g(\tau)$ of 
generation intervals (i.e., the time between when a person become 
infected and that person infects another person), the basic reproduction
number can be estimated via the Euler-Lotka equation:
\begin{equation}
1/\mathcal R_0 = \int \exp(-r\tau) g(\tau) d\tau.
\end{equation}
Therefore, early estimates of $\mathcal R_0$ must 
necessarily depend on the assumptions about $r$ and $g(\tau)$.

\begin{figure}[t]
\includegraphics[width=\textwidth]{comparison.pdf}
\caption{
Early estimates of $\mathcal R_0$ and associated assumptions about $r$ and $g(\tau)$.
}
\end{figure}

>>>>

\begin{flushleft}{
	\Large
	\textbf\newline{
		Evaluating uncertainties associated with early estimates of the basic reproduction number during the novel coronavirus (2019-nCoV) outbreak
	}
}
\newline
\\
Sang Woo Park\textsuperscript{1}
Jonathan Dushoff\textsuperscript{2,3}
\\

@@

\section{Introduction}

Since the end of December 2019, a novel coronavirus (2019-nCoV) continues to
spread in China and in other parts of the world.
As of January 27th, 2020, the World Health Organization has
confirmed 2798 cases, including 37 confirmed cases in 11 different
countries, outside China \citep{who27report}.
Although the virus is suspected to have originated
from animal reservoirs, a recent case from Viet Nam
demonstrated its ability to transmit between
humans \citep{who26report},
posing a greater threat for a wider spread.

Many researchers have been rushing to publish their 
analysis of the outbreak \citep{imaincov, riouncov,readncov,zhaoncov,majumderncov,liuncov} 
and, in particular, their
estimates of the basic 
reproductive number $\mathcal R_0$ (i.e., the 
average number of secondary cases generated 
by a primary case in a fully susceptible population).
The basic reproductive number is of particular interest 
because it allows prediction about the final size of an epidemic.
While their efforts are valuable, their analyses rely on several
assumptions that could immediately affect their estimates of $\mathcal R_0$ and
the associated uncertainties.

Here, we present a simple framework for evaluating uncertainties 
associated with parameter estimates
across a wide range of models.
Our results indicate that most published estimates of $\mathcal R_0$
are likely to be overly confident. 
We also lay down several principles that needs to be taken into
consideration.

@@

Early in an outbreak, $\mathcal R_0$ cannot be estimated directly.
Instead, estimates of $\mathcal R_0$ are often inferred from
the exponential growth rate $r$ can be estimated reliably from incidence data.
Given estimates of the exponential growth rate $r$ and the distribution $g(\tau)$ of 
generation intervals (i.e., the time between when a person become 
infected and that person infects another person), the basic reproduction
number can be estimated via the Euler-Lotka equation:
\begin{equation}
1/\mathcal R_0 = \int \exp(-r\tau) g(\tau) d\tau.
\end{equation}
In other words, estimates of $\mathcal R_0$
necessarily depends on the assumptions about 
growth rate $r$ and the shape of generation-interval distribution $g(\tau)$.

\begin{figure}[t]
\includegraphics[width=\textwidth]{compare_assumption.pdf}
\caption{
Early estimates of $\mathcal R_0$ and associated assumptions about $r$ and $g(\tau)$.
}
\end{figure}
